TensorAlloy: a highly efficient atomistic neural network program for alloys

### 2019-06-26 doc/papers/nn/manuscript.tex
CPC major improvements, revision 1.

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help. 
However, the traditional way to construct a 
machine learning model computation graph starts from pre-computed descriptors, 
but not positions, due to technical challenges. Thus, deriving analytical atomic 
force and virial stress of a machine learning model becomes quite difficult as 
the partial derivations of descriptors with respect to positions need separate 
implementations, which requires plenty of efforts and its efficiency can not be 
guaranteed.
In this paper, we designed a new method, named the virtual atom approach, for 
implementing the symmetry function descriptor and the corresponding atomistic 
neural networks. With the virtual atom approach, we are able to build 
computation graph from atomic positions \textemdash but not descriptors 
\textemdash to total energy directly, thus asking modern machine learning 
platforms to compute atomic force and virial stress becoming possible. We also 
derived a simple and machine learning friendly equation for calculating virial 
stress. All these new algorithms have been implemented in our Python package, 
TensorAlloy. TensorAlloy supports constructing machine learning interaction 
potentials for both \textit{molecules} and \textit{solids} and one hour is 
sufficient for TensorAlloy to obtain models with state-of-art level accuracy.
\end{abstract}

@@

In this work, we propose a new algorithm, named the virtual atom approach, to 
implement the symmetry function descriptor and the ANN model based on 
TensorFlow. This algorithm is very suitable for parallel execution. With this 
approach, stochastic gradient descent (SGD) based mini-batch training can be 
adopted. We also derive a general and vectorizable expression for computing 
virial stress for arbitrary ANN model. All these algorithms have been 
implemented in our Python program: \tensoralloy{}.

This paper is organized as follows. Section \ref{section:theory} introduces the 
theory of the symmetry function descriptor and the ANN framework. Section 
\ref{section:method} describes the virtual atom approach and a modified ANN 
model used by \tensoralloy{}. Section \ref{section:discussions} displays 
performances of \tensoralloy{} on two public datasets. The appendix 
\ref{section:appendix} gives derivation and implementation details of key 
algorithms. The \tensoralloy{} program can be obtained on request.

@@

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor, especially for alloys. The symmetry function descriptor 
consists of two sets of invariant (tranlational, rotational and permutational) 
functions: the radial symmetry function $\mathbf{G}^{(2)}$ and the angular 
symmetry function $\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, R_s) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, R_s)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2 = & \exp\left\{-\frac{\eta(\rijn - R_s)^2}{r_c^2} \right\}
\cdot f_c(\rijn) \\
g_4 = & \left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $R_s$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. $R_s$ may be fixed to zero \cite{AMP}. 
Equation \eqref{eq:fc_origin} can be further transformed to a vectorized 
expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

@@

\newcommand{\nmax}{
    N_{\mathrm{Ni}}^{\mathrm{max}}+N_{\mathrm{Mo}}^{\mathrm{max}}
}
\newcommand{\nijmax}{N_{ij}^{\mathrm{max}}}
\newcommand{\nijkmax}{N_{ijk}^{\mathrm{max}}}

In this section we will describe details of \tensoralloy{}: the virtual atom 
approach, the atomic residual potential and the loss function. 
The Ni-Mo binary alloy system will be used as the example to show this 
algorithm. The virtual atom approach is built upon TensorFlow. For simplicity, 
only radial symmetry function related algorithms are visualized in the figures.

\subsection{The virtual atom approach}
\label{section:virtual_atom_approach}

Fig \ref{fig:algo1} shows the initialization procedure of the virtual atom 
approach. The cutoff radius $r_{cut}$ must be fixed. The first step is to go 
through the entire dataset and determining four prerequisite constants: 
$N_{ij}^{\mathrm{max}}$, $N_{ijk}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$. 
Here $N_{ij}^{\mathrm{max}}$ represents the maximum number of neighbor pairs 
and $N_{ijk}^{\mathrm{max}}$ represents the maximum number of triples in 
arbitrary structure. 
$N_{ij}^{\mathrm{max}}$ and $N_{ijk}^{\mathrm{max}}$ only depend on 
$r_{cut}$. The associated codes are given in the appendix.
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$ indicate 
the maximum appearances of Ni and Mo in any structure, respectively. 
With $N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$, 
the global symbol list (GSL) can be created. GSL defines the global reference 
of this dataset and must be maintained during entire training. Then we can 
construct \textit{offsets}. This vector marks the starting indices of the 
\textbf{ordered} interaction pairs (e.g. Ni-Mo) and triples (e.g. Ni-Ni-Mo). 
The order of the interactions must be maintained as well.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{figures/Fig1-prepare.pdf}
\caption{\label{fig:algo1} The initialization step. 
\textbf{a)} describes the detections of $N_{\mathrm{Ni}}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$, $N_{ij}^{max}$, $N_{ijk}^{max}$ and the global
symbols list.
The \textit{offsets} of radial and angular functions and their corresponding 
symmetry function parameters are in \textbf{b)} and \textbf{c)}, respectively.
}
\end{figure}

Fig \ref{fig:algo2} shows the algorithm to construct the radial mapping matrix
$g_{\mathrm{map}}^2$ which plays a key role in building the computation graph 
(Fig \ref{fig:algo3}). This algorithm starts from the $N \times 3$ local 
positions matrix \textbf{R}. Here 'local' means original and 'global' means GSL. 
The first step is to find all neighbor pairs: $\mathbf{n}$, $i^{(0)}$ and 
$j^{(0)}$. This can be done by the 
\href{https://wiki.fysik.dtu.dk/ase/ase/neighborlist.html}{neighbor\_list} 
routine of ASE\cite{ase}. $\mathbf{n}$ is a $N_{ij} \times 3$ matrix 
representing the periodic boundary shift vectors where $N_{ij}$ is the total 
number of interaction pairs within $r_{cut}$.
Both $i^{(0)}$ and $j^{(0)}$ are $N_{ij}\times 1$ column vectors representing 
the indices of $i$ and $j$ of Equation \ref{eq:rijn}. 
$\mathbf{t}$ is a manually-created $N_{ij}\times 1$ column vector storing the 
types \textemdash represented by integers \textemdash of the corresponding 
interaction pairs.

The second step is to convert the local indices, $i^{(0)}$ and $j^{(0)}$, to 
global indices based on GSL and then \textbf{increase their values by one}. 
By padding with zeros, we can finally get the $N_{ij}^{\mathrm{max}} \times 3$ 
matrix $\mathbf{n_g}$ and $N_{ij}^{\mathrm{max}} \times 1$ column vectors 
$i^{(0)}_{\mathrm{+,g}}$, $j^{(0)}_{\mathrm{+,g}}$ and 
$\mathbf{t}_{\mathrm{g}}$. The increment is extremely important because the 
indices of all padded 'virtual' or 'dummy' atoms are fixed to 0.

The last step is to construct $g_{\mathrm{map}}^{(2)}$, which should be a 
$N_{ij}^{\mathrm{max}} \times 2$ matrix. Its first column is just 
$i^{(0)}_{\mathrm{+,g}}$ and its second column are corresponding offset values 
of $\mathbf{t}_{\mathrm{g}}$. $g_{\mathrm{map}}^{(2)}$, 
$j^{(0)}_{\mathrm{+,g}}$ and $\mathbf{n_g}$ can be pre-computed and cached. 
The algorithm to build $g_{\mathrm{map}}^{(4)}$ for angular symmetry functions 
is similar but too complicated to visualize. Thus, we provide a Python 
implementation in the appendix.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{figures/Fig2-v2gmap.pdf}
\caption{\label{fig:algo2} The second stage of the virtual atom approach: the 
construction of $g_{\mathrm{map}}^{(2)}$. Tensors marked red can be pre-computed 
and cached.}
\end{figure}

Fig \ref{fig:algo3} shows the final stage of the virtual atom approach: the 
computation graph for calculating radial symmetry function descriptors. This 
stage also starts from the $N \times 3$ local positions matrix \textbf{R} and 
involves two steps: the calculation of interatomic distances $r_{ij}$ 
(Fig \ref{fig:algo3}a) and the construction of $\mathbf{G}^{(2)}$ 
(Fig \ref{fig:algo3}b).

The initial step is to map local positions \textbf{R} to
$(N_{\mathrm{Ni}}^{\mathrm{max}}+N_{\mathrm{Mo}}^{\mathrm{max}}+1) \times 3$
global positions \textbf{R'} based on GSL. The first row (marked blue) of 
$\mathbf{R'}$ are all zeros representing 'positions' of the virtual atom. 
Then with 
\href{https://www.tensorflow.org/versions/r1.12/api_docs/python/tf/gather}
{tf.gather}, the global indices $i^{(0)}_{\mathrm{+,g}}$ and 
$j^{(0)}_{\mathrm{+,g}}$ of Fig \ref{fig:algo2} can be transformed to global 
positions: $\mathbf{R'_i}$ and $\mathbf{R'_j}$. Thus, the interatomic distances 
$r_{ij}$ can be simply calculated with Equation \ref{eq:rijn}.

Now it's time to get radial symmetry function descriptors matrix 
$\mathbf{G}^{(2)}$. 
As shown in Fig \ref{fig:algo3}b, this step can be executed in parallel: 
for each $\eta_i$, we first calcualte $g_2(r_{ij}, \eta_i)$ with 
Equation \ref{eq:g2} and then map the $N_{ij}^{\mathrm{max}}$-length vector 
$g_2(r_{ij}, \eta_i)$ to 
$(N_{\mathrm{Ni}}^{\mathrm{max}}+N_{\mathrm{Mo}}^{\mathrm{max}}+1) \times 16$ 
matrix $\mathbf{G}^{(2)}_{\eta_i}$ with $g^{\eta_i}_{\mathrm{map}}$ and 
\href{https://www.tensorflow.org/versions/r1.12/api_docs/python/tf/scatter_nd}
{tf.scatter\_nd}. Here $16$ is the multiplacation of 4 ($\eta$) and 4 
(radial interactions). 
$g^{\eta_i}_{\mathrm{map}}$ is calculated by adding a constant
$i - 1$ to all values of the second column of $g^{(2)}_{\mathrm{map}}$. Only 
shaded columns in Fig \ref{fig:algo3}b have non-zero values.
In fact, the first column of $g^{\eta_i}_{\mathrm{map}}$ means the indices of 
the central atoms in Equation \ref{eq:G2} (or rows of $\mathbf{G}^{(2)}$) and 
the second column marks the corresponding $\eta_i$ in Fig \ref{fig:algo1} (or 
columns of $\mathbf{G}^{(2)}$).
The first row of each $\mathbf{G}^{(2)}_{\eta_i}$ reprsents the descriptors of 
the virtual atom. Removing these rows with
\href{https://www.tensorflow.org/versions/r1.12/api_docs/python/tf/split}
{tf.split}, we get $\hat{\mathbf{G}}^{(2)}_{\eta_i}$.
The final radial symmetry function descriptors, $\mathbf{G}^{(2)}$ is just the 
sum of all $\hat{\mathbf{G}}^{(2)}_{\eta_i}$.

\begin{figure*}[h!]
\centering
\includegraphics[scale=0.5]{figures/Fig3-G2.pdf}
\caption{\label{fig:algo3} 
\textbf{a)} The calculation of the interatomic distances $r_{ij}$.
\textbf{b)} The construction of the radial symmetry function descriptors 
$\mathbf{G}^{(2)}_{\eta_i}$. Rows that are marked blue correspond to the virtual 
atom and will be discared before feeding $\mathbf{G}$ into atomistic neural 
networks.
}
\end{figure*}

\subsection{The data flow}

The following equation briefly demonstrates the core data flow of 
\textbf{TensorAlloy} during a mini-batch training: 
\begin{equation}
    \mathbf{R'_{b}} \rightarrow 
    \mathbf{G_b} \rightarrow 
    \mathbf{E_b} \rightarrow
    \mathbf{F_b} \rightarrow 
    \mathbf{\epsilon_b} \rightarrow
    \mathbf{Loss}
\end{equation}
where the $N_{b} \times (\nmax + 1) \times 3$ matrix $\mathbf{R'_b}$ is a batch 
of positions matrices ($N_b \ge 1$ is the batch size), $\mathbf{G_b}$ is the 
corresponding symmetry function descriptors matrix of shape 
$N_{b} \times (\nmax) \times 16$ and $\mathbf{E_b}$ represents 
total energies of the structures. Section \ref{section:virtual_atom_approach} 
describes the implementation of $\mathbf{R'_{b}} \rightarrow \mathbf{G_b}$. 
To implement $\mathbf{G_b} \rightarrow \mathbf{E_b}$, \textbf{TensorAlloy} uses 
1D-convolutional neural networks with $1 \times 1$ kernels\cite{kCON}. The total 
forces $\mathbf{F_{b}}$ and the virial stress $\mathbf{\epsilon_{b}}$ are 
calculated with Equation \ref{eq:dEdr} and Equation \ref{eq:stress} directly.

\subsection{The residual model}

To fit the total energy, we slightly modified the total energy expression of 
Equation \ref{eq:general_e_total}:
\begin{align}
\label{eq:e_total_res}
E^{total} & = E^{residual} + E^{static} \nonumber \\
& = \sum_{i=1}^{N}{
    \mathbf{NN}_{el}\left( \left\{ G_i \right\} \right)
} + \sum_{el}{n_{el}E_{el}}
\end{align}
where $\{E_{el}\}$ is a set of \textit{trainable} scalar variables representing 
the \textit{static} energy of each type of element\cite{kCON}. The NN in 
Equation \ref{eq:e_total_res} just describes the atomistic interactions 
(the \textit{residual} energy). The 
introduction of $E^{static}$ can significantly limit value range of NN outputs, 
leading to faster convergence speed and higher training stability 
(Fig \ref{fig:benchmark_qm7}c). The initial $\{E_{el}\}$ are calculated by 
solving the linear system $Ax=b$ where $A$ is a $N_{data} \times N_{el}$ matrix 
and $b$ is a $N_{data}\times 1$ vector. $A(i,j)$ is the number of $j^{th}$
element in structure $i$ and $b(i)$ is the total energy of structure $i$.

>>>>

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help. 
However, the traditional way to construct a 
machine learning model computation graph starts from pre-computed descriptors, 
but not positions, due to technical challenges. Thus, deriving analytical atomic 
force and virial stress of a machine learning model becomes quite difficult as 
the partial derivations of descriptors with respect to positions need separate 
implementations, which requires plenty of efforts and its efficiency can not be 
guaranteed.
In this paper, we designed a new method, named the virtual atom approach, for 
implementing the symmetry function descriptor and the corresponding atomistic 
neural networks. With the virtual atom approach, we are able to build 
computation graph from atomic positions \textemdash but not descriptors 
\textemdash to total energy directly, thus asking modern machine learning 
platforms to compute atomic force and virial stress becoming possible. We also 
derived a simple and machine learning friendly equation for calculating virial 
stress. All these new algorithms have been implemented in our Python package, 
TensorAlloy. TensorAlloy supports constructing machine learning interaction 
potentials for both \textit{molecules} and \textit{solids}.
\end{abstract}

@@

In this work, we propose a new algorithm, named the virtual atom approach, to 
implement the symmetry function descriptor and the ANN model based on 
TensorFlow. This algorithm is very suitable for parallel execution. With this 
approach, stochastic gradient descent (SGD) based mini-batch training can be 
adopted. We also derive a general and vectorizable expression for computing 
virial stress for arbitrary ANN model. All these algorithms have been 
implemented in our Python program: TensorAlloy.

This paper is organized as follows. Section \ref{section:theory} introduces the 
theory of the symmetry function descriptor and the ANN framework. Section 
\ref{section:method} describes the virtual atom approach and a modified ANN 
model used by TensorAlloy. Section \ref{section:discussions} displays 
performances of TensorAlloy on two public datasets. The appendix 
\ref{section:appendix} gives derivation and implementation details of key 
algorithms. Implementations of the core algorithms and some trained models can 
be obtained from GitHub (https://github.com/Bismarrck/vap) freely.

@@

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor, especially for alloys. The symmetry function descriptor 
consists of two sets of invariant (tranlational, rotational and permutational) 
functions: the radial symmetry function $\mathbf{G}^{(2)}$ and the angular 
symmetry function $\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, \omega) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, \omega)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2(\eta, i, j, \omega) = 
& \exp\left\{-\frac{\eta(\rijn - \omega)^2}{r_c^2} \right\} \cdot f_c(\rijn) \\
\label{eq:g4}
g_4(
    \beta, \gamma, \zeta, 
    i, j, k, 
    \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3) = & 
\left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
\nonumber \\
& \times f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $\omega$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. Equation \eqref{eq:fc_origin} can be further 
transformed to a vectorized expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

@@

\newcommand{\nmax}{
    N_{\mathrm{Ni}}^{\mathrm{max}}+N_{\mathrm{Mo}}^{\mathrm{max}}
}
\newcommand{\nijmax}{N_{ij}^{\mathrm{max}}}
\newcommand{\nijkmax}{N_{ijk}^{\mathrm{max}}}
\newcommand{\nelmax}{N^{\mathrm{max}}_{\mathrm{el}}}
\newcommand{\nel}{N_{\mathrm{element}}}
\newcommand{\nvap}{N^{\mathrm{vap}}}
\newcommand{\nnimax}{N^{\mathrm{max}}_{\mathrm{Ni}}}
\newcommand{\nmomax}{N^{\mathrm{max}}_{\mathrm{Mo}}}

In this section we will describe the program design and the core algorithm: 
the virtual atom approach (VAP). The most noticable feature of the TensorAlloy 
program is that the calculation of symmetry function descriptors is finally 
implemented in the compuation graph with the help of VAP. Thus, we can utilize 
the \textbf{AutoGrad} routine of TensorFlow to derive atomic forces and virial 
stress automatically and also efficiently.

The Ni-Mo binary alloy system will be used as the example to show this 
algorithm. For simplicity, only radial symmetry function related algorithms are 
visualized in the figures. The associated codes (Python3.7) are given in the 
appendix. A demo script can be downloaded from GitHub freely 
(https://github.com/Bismarrck/vap).

Fig \ref{fig:program_design} shows the general design of TensorAlloy.
TensorAlloy has two phases: 1) is the training phase and 2) represents the 
prediction phase. These two phases have slightly different focusing points:

In the training phase, efficiency is the major concern. To accelerate model 
training, batching (or mini-batch training) is one of the most commonly used 
techniques. Batching requires aligned inputs. However, hundreds of different 
stoichiometries may exist in one dataset. To batch such structures, an universal 
approach of expressing different structures in a single reference system (the 
global symbol list or GSL) is necessary.

In the prediction phase, things may be a bit different. The stoichiometry of 
the input structure may not be included in the training dataset and it can vary 
significantly. Taking the example of Ni-Mo, the trained model must be capable of 
predicting any $\mathrm{Ni}_{x}\mathrm{Mo}_{y}$ solid while $x$ and $y$ can be 
either as large as 50000 or just zero. Hence, VAP must be flexible enough to 
handle these situations.

In the following subsection, we will discuss the details of VAP.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.58]{figures/Fig1.pdf}
\caption{\label{fig:program_design} The general design of TensorAlloy. 
\textbf{(1)} demonstrates the training phase while \textbf{(2)} represents the 
prediction phase.
}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Preparation}
\label{section:preparation}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.2
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{figures/Fig2.pdf}
\caption{\label{fig:Fig2} The initialization step. 
\textbf{(1)} describes the detections of $N_{\mathrm{Ni}}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and the global symbol list (GSL). The red 'X' 
represents the inserted virtual atom.
\textbf{(2)} shows the detections of $\nijmax$ and $\nijkmax$ given $r_{cut}$.
}
\end{figure}

Fig \ref{fig:Fig2} shows the initialization of VAP. The cutoff radius $r_{cut}$ 
must be fixed. The initialization involves two steps. The first step is to build 
the global symbol list (GSL), which acts as the universal reference system. A 
virtual atom is inserted at the very first position of this list. The benefits 
will be discussed later. To build GSL, all $\nelmax$ must be obtained. 
$\nelmax$ indicates the maximum appearances of an element in any structure and 
$\nel$ represents the total number of unique chemical elements:
\begin{equation}
\label{eq:n_vap}
\nvap = 1 + \sum_{\mathrm{el}}{\nelmax}
\end{equation}
For the Ni-Mo dataset, $\mathrm{el} \in [\mathrm{Ni}, \mathrm{Mo}]$ and 
$\nel = 2$. 
Table \ref{table:GSL} gives some examples of the GSL mapping. 'X' denotes the 
inserted virtual atom and Ni$_{108(54)}$ means the first 54 Ni atoms are mapped 
from the local stoichiometry. In the training phase, all $\nelmax$ are constant.
In the prediction phase, $\nelmax$ shall be determined dynamically. One should 
note that the minimum value of $\nelmax$ should be 1 due to technical reason (so 
that the computation graph of TensorFlow can be properly executed). The codes to
get $\nelmax$ and GSL are given in Appendix(C). With GSL setup, atomic positions
and forces can be transformed as well. Appendix(D) gives the implementation of 
the virtual-atom mapping mechanism.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccccc}
\hline
 & Stoichiometry & $\nnimax$ & $\nmomax$ & $N^{\mathrm{vap}}$ 
 & Stoichiometry (GSL) \\
\hline
 & Ni$_{54}$ &  &  &  & XNi$_{108(54)}$Mo$_{176(0)}$ \\
Training & Ni$_{54}$Mo$_{12}$ & 108 & 176 & 285 
 & XNi$_{108(54)}$Mo$_{176(12)}$ \\
 & Mo$_{12}$ &  &  &  & XNi$_{108(0)}$Mo$_{176(12)}$ \\
\hline
 & Ni$_{54}$ & 54 & 1 & 56 & XNi$_{54(54)}$Mo$_{1(0)}$ \\
Prediction & Ni$_{54}$Mo$_{12}$ & 54 & 12 & 68 & XNi$_{54(54)}$Mo$_{12(12)}$ \\
 & Mo$_{12}$ & 1 & 12 & 14 & XNi$_{1(0)}$Mo$_{12(12)}$ \\
\hline
\end{tabular}
\caption{\label{table:GSL}
The mappings of local stoichiometries to their GSL forms, taking examples of the 
SNAP/Ni-Mo dataset.
}
\end{table}

Then we can determine another two key constants: $\nijmax$ and $\nijkmax$. 
Here $N_{ij}^{\mathrm{max}}$ represents the maximum number of neighbor pairs in
arbitrary structure and the subscript 'ijk' denotes triples. 
$N_{ij}^{\mathrm{max}}$ and $N_{ijk}^{\mathrm{max}}$ only depend on 
$r_{cut}$ and the associated codes are given in Appendix(B). 
Table \ref{table:nij_nijk_max} summarizes $\nijmax$ and $\nijkmax$ at different
$r_{cut}$ of the two tested datasets: QM7 and SNAP/Ni-Mo.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table.2
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccc}
\hline
 Dataset & $r_{cut}$(\AA) & $\nijmax$ & $\nijkmax$ \\
 QM7 & 6.5 & 506 & 5313 \\
\hline
 SNAP/Ni-Mo &  4.6 & 7200 & 176400 \\
 & 6.0 & 12384 & 526320 \\
 & 6.5 & 16284 & 965338 \\
\hline
\hline
\end{tabular}
\caption{\label{table:nij_nijk_max}
$\nijmax$ and $\nijkmax$ of SNAP/Ni-Mo and QM7 at different $r_{cut}$.
}
\end{table}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.B
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Interatomic distances}
\label{section:interatomic_distance}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.3
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{figures/Fig3.pdf}
\caption{\label{fig:Fig3} The pseudo computation graph with execution orders for 
calculating interatomic distances $\mathbf{r}$ from atomic positions 
$\mathbf{R}$ and lattice tensor $\mathbf{h}$.
}
\end{figure}

Both the radial (Equation \ref{eq:g2}) and the angular (Equation \ref{eq:g4}) 
functions only interact with interatomic distances. In order to compute symmetry
function descriptors, local atomic positions shall be transformed to GSL 
interatomic distances first. Fig \ref{fig:Fig3} demonstrates the pseudo 
computation graph for calculating interatomic distances 
$\mathbf{r}^{\mathrm{GSL}}$ for radial descriptors:

\begin{itemize}

    \item[1.] 
    Find all (i,j) neighbor pairs using the routine 
    \href{https://wiki.fysik.dtu.dk/ase/ase/neighborlist.html}
    {\textit{neighbor\_list}} implemented in ASE\cite{ase}. $\mathbf{R}$ 
    represents local atomic positions, $\mathbf{h}$ is the lattice tensor. 
    $\mathbf{i}$ and $\mathbf{j}$ are vectors of length $N_{\mathrm{ij}}$ and
    $\mathbf{n}$ is the $N_{\mathrm{ij}} \times 3$ periodic boundary shift 
    matrix. In the training phase $N_{\mathrm{ij}} \le \nijmax$; in the 
    prediction phase, $\nijmax$ is exactly equal to $N_{\mathrm{ij}}$.

    \item[2.] 
    Map $\mathbf{R}$, $\mathbf{i}$, $\mathbf{j}$ and $\mathbf{n}$ to
    their corresponding GSL forms with padding zeros. One should note that these 
    GSL arrays can be cached for training.

    \item[3.] 
    Broadcast $\mathbf{R}^{\mathrm{GSL}}$ to $\mathbf{R}_{i}^{\mathrm{GSL}}$ and
    $\mathbf{R}_{j}^{\mathrm{GSL}}$ with $\mathbf{i}^{\mathrm{GSL}}$ and 
    $\mathbf{j}^{\mathrm{GSL}}$ respectively.

    \item[4.]
    Compute the displacements in the original cell:
    $\mathbf{D}_{ij} = \mathbf{j}^{\mathrm{GSL}} - \mathbf{i}^{\mathrm{GSL}}$

    \item[5.]
    Compute the total periodic displacements:
    $\mathbf{D}_{pbc} = \mathbf{n}^{\mathrm{GSL}} \mathbf{h}$

    \item[6.]
    Compute the overall GSL displacements:
    $\mathbf{D}^{\mathrm{GSL}} = \mathbf{D}_{pbc} + \mathbf{D}_{ij}$

    \item[7.] 
    Compute the interatomic distances $\mathbf{r}^{\mathrm{GSL}}$.

\end{itemize}

Steps 1-2 are implemented in the function \textit{get\_g2\_map} in Appendix(E).
Steps 3-7 are implemented in the method \textit{get\_rij} of class 
\textit{SymmetryFunction} in Appendix(G).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Radial descriptors}
\label{section:radial_descriptors}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.4
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

Now we can start the last step: computing radial symmetry descriptors 
$\mathbf{G}^{(2)}$ from $\mathbf{r}^{\mathrm{GSL}}$. For simplicity, we will use 
the prediction phase to demonstrate this final section.

Suppose we have $N_{\tau}$ different $(\eta, \omega)$ pairs for Equation 
\ref{eq:G2} and $(\eta, \omega)_{\tau} (0 \le \tau < N_{\tau})$ represents the 
$(\tau+1)$-th pair. In the prediction phase, $\mathbf{r}^{\mathrm{GSL}}$ should 
be a vector of length $\nijmax$ and the desired output $\mathbf{G}^{(2)}$ will 
be a matrix of shape $[\nvap, (\nel)^2 \cdot N_{\tau}]$, as shown in 
Fig \ref{fig:Fig5}\textbf{d}:
\begin{itemize}
    \item[1.] 
    The total number of rows is $\nvap$ (Equation \ref{eq:n_vap}) and each row
    represents the complete radial symmetry function descriptors of an atom. The 
    first row always corresponds to the inserted virtual atom.
    \item[2.]
    The total number of columns is $\nel \cdot \nel \cdot N_{\tau}$. Here 
    $\nel \cdot \nel$ indicates the number of element-element interaction types 
    and each interaction is described by $N_{\tau}$ feature values.
\end{itemize}
